Make manuscript figures

### 2022-06-14 scripts/R/make_manuscript_figures.rmd
add code to summarize FPR

Load data
```{r}
accuracy_simulation_results <- read.delim("output/2021-08-31_estimator_accuracy_MLE.txt")
tpr_simulation_results <- read.delim("output/2021-08-31_GWAS_TPR_MLE.txt")
example_simulation_results <- read.delim("output_revisions/2022-06-11_simulateFullDataset.txt")
tree <- ape::read.tree(file = paste(OUTDIR, "pathogen_no_outgroup.newick", sep = "/"))
tree_arboricola <- ape::read.tree(file = paste(OUTDIR_XANTHAMONAS, "pathogen_nj.newick", sep = "/"))
traits <- read.csv(file = paste(OUTDIR, "poumm_corrected_traits.csv", sep = "/"))
traits_arboricola <- read.csv(file = paste(OUTDIR_XANTHAMONAS, "poumm_corrected_traits.csv", sep = "/"))
tree_w_trait_data <- treeio::full_join(tree, traits, by = "label")
tree_w_trait_data_arboricola <- treeio::full_join(tree_arboricola, traits_arboricola, by = "label")
POUMM_fit <- get(load(file = paste(OUTDIR, "POUMM_fit.RData", sep = "/")))
POUMM_fit_arboricola <- get(load(file = paste(OUTDIR_XANTHAMONAS, "POUMM_fit.RData", sep = "/")))
poumm_parameter_estimates <- read.csv(file = paste(OUTDIR, "poumm_parameter_estimates.csv", sep = "/"))
poumm_parameter_estimates_arboricola <- read.csv(file = paste(OUTDIR_XANTHAMONAS, "poumm_parameter_estimates.csv", sep = "/"))
pc_scores <- read.csv(file = paste(OUTDIR, "shcs_vs_hapmap_pc_scores.csv", sep = "/"))
gwas_results <- read.delim(file = paste(OUTDIR, "gwas_results.pvals.effectsize.txt", sep = "/"), sep = "", header = T)
gwas_results_xanthomonas <- read.delim(file = paste(OUTDIR_XANTHAMONAS, "gwas_results.pvals.effectsize.txt", sep = "/"), sep = "", header = T)
spvl_values <- read.csv(file = paste(OUTDIR, "spvl_values.csv", sep = "/"))
```

@@

# Calculate TPR for each simulation
sig.level <- correctSignificanceLevelForMultTesting(
        significance.level = 0.05,
        significance.correction = "Bonferroni",
        N.tests = 20)
data$TP <- ifelse(data$p.value < sig.level & data$is.causal.variant, T, F)
data$FP <- ifelse(data$p.value < sig.level & !(data$is.causal.variant), T, F)
data.TPR <- as.data.frame(
        data %>%
                dplyr::group_by(runID, inference.method, alpha, sigma, H2, var.z, p, delta) %>%
                dplyr::summarise(
                        TPR = sum(TP, na.rm = T)/sum(is.causal.variant),
                        FP = sum(FP), K = dplyr::n(),
                        N.noncausal.variants = sum(!(is.causal.variant))))

@@

@@

Estimated trait values Xanthomonas
```{r}
p1 <- ggplot(data = traits_arboricola, aes(x = trait, y = h.MWA)) +
      geom_point() +
      shared_theme +
      labs(x = "Calculated QDR trait", y = ehat_name)
p2 <- ggExtra::ggMarginal(p1, type="histogram")
ggsave(plot = p2, filename = paste(FIGDIR, "trait_values_arboricola.png", sep = "/"),
      width = linewidth * 2/3, height = height * 2/3, units = "cm")
```
Compare variance-covariance matrix under POUMM
```{r}
# Get variance-covariance matrix from POUMM
corr_HIV <- POUMM::covVTipsGivenTreePOUMM(
        tree = tree, 
        alpha = poumm_parameter_estimates[[which(poumm_parameter_estimates$stat == 'alpha'), 'PostMean']], 
        sigma = poumm_parameter_estimates[[which(poumm_parameter_estimates$stat == 'sigma'), 'PostMean']], 
        sigmae = poumm_parameter_estimates[[which(poumm_parameter_estimates$stat == 'sigmae'), 'PostMean']],
        corr = T
)
corr_arboricola <- POUMM::covVTipsGivenTreePOUMM(
        tree = tree_arboricola, 
        alpha = poumm_parameter_estimates_arboricola[[which(poumm_parameter_estimates_arboricola$stat == 'alpha'), 'PostMean']], 
        sigma = poumm_parameter_estimates_arboricola[[which(poumm_parameter_estimates_arboricola$stat == 'sigma'), 'PostMean']], 
        sigmae = poumm_parameter_estimates_arboricola[[which(poumm_parameter_estimates_arboricola$stat == 'sigmae'), 'PostMean']],
        corr = T
)
max(corr_HIV[upper.tri(corr_HIV)])
min(corr_HIV[upper.tri(corr_HIV)])
max(corr_arboricola[upper.tri(corr_arboricola)])
min(corr_arboricola[upper.tri(corr_arboricola)])
```

>>>>

Load data
```{r}
accuracy_simulation_results <- read.delim("output/2021-08-31_estimator_accuracy_MLE.txt")
tpr_simulation_results_old <- read.delim("output/2021-08-31_GWAS_TPR_MLE.txt")
tpr_simulation_results <- read.delim("output_revisions/2022-06-13_GWAS_TPR_MLE.txt")
example_simulation_results <- read.delim("output_revisions/2022-06-11_simulateFullDataset.txt")
tree <- ape::read.tree(file = paste(OUTDIR, "pathogen_no_outgroup.newick", sep = "/"))
tree_arboricola <- ape::read.tree(file = paste(OUTDIR_XANTHAMONAS, "pathogen_nj.newick", sep = "/"))
traits <- read.csv(file = paste(OUTDIR, "poumm_corrected_traits.csv", sep = "/"))
traits_arboricola <- read.csv(file = paste(OUTDIR_XANTHAMONAS, "poumm_corrected_traits.csv", sep = "/"))
tree_w_trait_data <- treeio::full_join(tree, traits, by = "label")
tree_w_trait_data_arboricola <- treeio::full_join(tree_arboricola, traits_arboricola, by = "label")
POUMM_fit <- get(load(file = paste(OUTDIR, "POUMM_fit.RData", sep = "/")))
POUMM_fit_arboricola <- get(load(file = paste(OUTDIR_XANTHAMONAS, "POUMM_fit.RData", sep = "/")))
poumm_parameter_estimates <- read.csv(file = paste(OUTDIR, "poumm_parameter_estimates.csv", sep = "/"))
poumm_parameter_estimates_arboricola <- read.csv(file = paste(OUTDIR_XANTHAMONAS, "poumm_parameter_estimates.csv", sep = "/"))
pc_scores <- read.csv(file = paste(OUTDIR, "shcs_vs_hapmap_pc_scores.csv", sep = "/"))
gwas_results <- read.delim(file = paste(OUTDIR, "gwas_results.pvals.effectsize.txt", sep = "/"), sep = "", header = T)
spvl_values <- read.csv(file = paste(OUTDIR, "spvl_values.csv", sep = "/"))
```

@@

# Calculate TPR for each simulation
sig.level <- correctSignificanceLevelForMultTesting(
        significance.level = 0.05,
        significance.correction = "Bonferroni",
        N.tests = 100)
data$TP <- ifelse(data$p.value < sig.level & data$is.causal.variant, T, F)
data$FP <- ifelse(data$p.value < sig.level & !(data$is.causal.variant), T, F)
data.TPR <- as.data.frame(
        data %>%
                dplyr::group_by(runID, inference.method, alpha, sigma, H2, var.z, p, delta) %>%
                dplyr::summarise(
                        TPR = sum(TP, na.rm = T)/sum(is.causal.variant),
                        FP = sum(FP), K = dplyr::n(),
                        N.noncausal.variants = sum(!(is.causal.variant))))

@@

fp_summary <- data %>%
        group_by(inference.method) %>%
        summarise(FPR = sum(FP, na.rm = T) / sum(!is.na(FP)), n_non_na = sum(!is.na(FP)))

@@

Estimated trait values Xanthomonas
```{r}
p1 <- ggplot(data = traits_arboricola, aes(x = trait, y = h.MWA)) +
      geom_point() +
      shared_theme +
      labs(x = "Calculated QDR trait", y = ehat_name)
p2 <- ggExtra::ggMarginal(p1, type="histogram")
ggsave(plot = p2, filename = paste(FIGDIR, "trait_values_arboricola.png", sep = "/"),
      width = linewidth * 2/3, height = height * 2/3, units = "cm")
```
Compare variance-covariance matrix under POUMM
```{r}
# Get variance-covariance matrix from POUMM
corr_HIV <- POUMM::covVTipsGivenTreePOUMM(
        tree = tree, 
        alpha = poumm_parameter_estimates[[which(poumm_parameter_estimates$stat == 'alpha'), 'PostMean']], 
        sigma = poumm_parameter_estimates[[which(poumm_parameter_estimates$stat == 'sigma'), 'PostMean']], 
        sigmae = poumm_parameter_estimates[[which(poumm_parameter_estimates$stat == 'sigmae'), 'PostMean']],
        corr = T
)
corr_arboricola <- POUMM::covVTipsGivenTreePOUMM(
        tree = tree_arboricola, 
        alpha = poumm_parameter_estimates_arboricola[[which(poumm_parameter_estimates_arboricola$stat == 'alpha'), 'PostMean']], 
        sigma = poumm_parameter_estimates_arboricola[[which(poumm_parameter_estimates_arboricola$stat == 'sigma'), 'PostMean']], 
        sigmae = poumm_parameter_estimates_arboricola[[which(poumm_parameter_estimates_arboricola$stat == 'sigmae'), 'PostMean']],
        corr = T
)
max(corr_HIV[upper.tri(corr_HIV)])
min(corr_HIV[upper.tri(corr_HIV)])
max(corr_arboricola[upper.tri(corr_arboricola)])
min(corr_arboricola[upper.tri(corr_arboricola)])
```
Look for false positives in the simulated data
```{r}
sig.level <- correctSignificanceLevelForMultTesting(
        significance.level = 0.05,
        significance.correction = "Bonferroni",
        N.tests = 100)

example_simulation_results <- example_simulation_results %>%
        mutate(
                FP <- ifelse(p.value < sig.level & !(is.causal.variant), T, F)
        )
```
Compare other ML phylogenies for HIV data
```{r}
trees <- ape::read.tree("output_revisions/pathogen.treels.head500")

# Calculate tree distances
tree_dists <- dist.topo(trees, method = "PH85")
```
